---
title: "NBHF Classes"
author: Jackson Vanfleet-Brown
date: 03/06/2024
format: html
editor: source
echo: false
toc: true
---
```{r}
#| output: false

library(identidrift)
library(tidyverse)
library(PAMpal)
library(knitr)
```

## NBHF Classes

```{r}
#| label: tbl-ndets
#| tbl-cap: "Size of each class, represented by number of distinct detections and number of events over which the detections occurred, for each species class in the training set."
#| 
matrix(data = c(sapply(train, nDetections, distinct=TRUE),
                sapply(train, \(x) length(events(x)))),
       nrow=3,
       ncol=2,
       dimnames = list(c("*Kogia*", "Dall's porpoise",
                         "harbor porpoise"),
                       c("detections","events"))) %>%
  kable()
```

### Results

I can add some general remarks here about broad differences in @fig-avgspec the shapes of the different profiles and what features led us to select them as high-quality clicks. *Kogia* in @fig-avgspec-ks, Dall's in @fig-avgspec-pd, and harbor in @fig-avgspec-pp.

```{r}
classPlot <- function(species, ylim = c(-25, 0)) {
  par(mfrow=c(1,2))
  calculateAverageSpectra(train[[species]],
                          evNum=1:length(events(train[[species]])),
                          sort=TRUE, showBreaks = FALSE, title="",
                          filterfrom_khz = 100, filterto_khz = 160,
                          ylim = ylim, flim=c(100000,160000),
                          norm=TRUE)
} 
```


::: {#fig-avgspec layout-nrow=3}

```{r #fig-avgspec-ks}
#| fig-cap: "*Kogia* spp."
#| cache: true
classPlot("ks")
```

```{r #fig-avgspec-pd}
#| fig-cap: "Dall's porpoise"
#| cache: true
classPlot("pd")
```

```{r #fig-avgspec-pp}
#| fig-cap: "Harbor porpoise"
#| cache: true
classPlot("pp")
```

Averaged spectra and concatenated spectrograms for all species classes in the training set.
:::

## Table of summary statistics
```{r}
#| label: tbl-class-stats
#| tbl-cap: "Summary statistics of acoustic features for each species class. Mean (sd) is given in the top row of each cell, and median (min - max) is given in the bottom row of each cell."
#| warning: false
#| error: false
#| message: false

varsWanted <- c("BW_3dB", "BW_10dB", "centerkHz_3dB", "centerkHz_10dB", "duration",
                "fmax_10dB", "fmax_3dB", "fmin_3dB", "fmin_10dB", "peak", "Q_3dB",   
                 "Q_10dB", "peakTime")

train.ec %>%
  select(species, all_of(varsWanted)) %>%
  pivot_longer(cols=-species, names_to = "var") %>%
  group_by(species, var) %>%
  summarize(mean=mean(value), sd=sd(value), median=median(value),
            min=min(value), max=max(value)) %>%
  mutate_if(is.numeric, round, 2) %>%
  mutate(row1=str_c(mean, " (", sd, ")")) %>%
  mutate(row2=str_c(median, " (", min, " - ", max, ")")) %>% 
  mutate(entry=str_c(row1, row2, sep="<br>")) %>%
  select(species, var, entry) %>%
  pivot_wider(names_from = var, values_from = entry) %>%
  mutate(species = case_match(species, "ks" ~ "*Kogia* spp.",
                              "pd" ~ "Dall's porpoise",
                              "pp"~ "harbor porpoise")) %>% 
  column_to_rownames("species") %>%
  t() %>%
  kable(align="c")
```

### Discussion points

This is important for those who are interested in seeing the exact profiles of these types of signals and could be helpful for people studying NBHF clicks in other regions, to compare to what we consider to be a high-quality click.